---
title: "ESM204_HW3"
author: "Nickolas McManus"
date: "4/15/2022"
output: 
  html_document: 
    toc: yes
    toc_float: yes
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)
library(janitor)
library(equatiomatic)
library(broom)
library(rootSolve)
```

```{r}
# read in data
e_data <-read_csv(here("HW3_data.csv")) %>% 
  
  #make names better and get rid of worthless first column
  clean_names() %>% 
  select(-1)
```


```{r}
# Linear regression models

model_demand_l <- lm(price_cents ~ q_low_kwh, data = e_data)
model_demand_h <- lm(price_cents ~ q_high_kwh, data = e_data)

low_tidy <- tidy(model_demand_l)
low_tidy

high_tidy <- tidy(model_demand_h)
high_tidy
```

need to rearrange the parameter to get Q(P)! 

Qgg = Qlow(P) + Qlow(h) 

Importantly, since they-intercepts are different, we know that Qagg(P) will have a kink. I include an ifelse() statement to take care of the kink.

define a function to get demand

```{r}
# Importantly, since they-intercepts are different, we know that Qagg(P) will have a kink. I include an ifelse() statement to take care of the kink.

# define a function to get demand
demand <- function(p, model){
  q <- (p - model$coefficients[[1]])/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}

# for each p level, return aggregate demand
demand_agg <- function(p){
  q <- demand(p, model_demand_l) + demand(p, model_demand_h)
  return(q)
}

# trying to replicate for low-income demand
demand_low <- function(p){
  q <- demand(p, model_demand_l)
  return(q)
}

# trying to replicate for high-income demand
demand_high <- function(p){
  q <- demand(p, model_demand_h)
  return(q)
}

```

```{r}
# Sandy's code for Qagg
price = seq(0, 30, length.out = 100)
Qagg <- map(price, demand_agg) %>% unlist()

df<- tibble(Qagg = Qagg, price = price)

# trying to do the same for Qlow
Qlow <- map(price, demand_low) %>% unlist()
Qlow_df <- tibble(Qlow = Qlow, price = price)

# trying to do same for Qhigh
Qhigh <- map(price, demand_high) %>% unlist()
Qhigh_df <- tibble(Qhigh = Qhigh, price = price)

Qhigher <- map(price, demand_high) %>% unlist()
Qhigher_df <- tibble (Qhigher = Qhigher, price = price)


ggplot() +
  # graph the agg demand
  geom_line(data = df, aes(Qagg, price)) +
  
  # graph low demand??
  geom_line(data = Qlow_df, aes(Qlow, price)) +
  
  # graph high demand??
  geom_line(data = Qhigh_df, aes(Qhigh, price)) +
  
  # graph the low income points
  geom_point(data = e_data, aes(x = q_low_kwh, y = price_cents), color = "slateblue", alpha = 0.5, size = 2) +
  # graph the high income points
  geom_point(data = e_data, aes(x = q_high_kwh, y = price_cents), color = "sienna", alpha = 0.5, size = 2) +

    labs(x = "\nQuantity (kWh)",
       y = "Price ($/kWh)\n") +
  
  theme_classic() +
  
  theme(
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold")) +
  
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0))
  

```


<br>

### Q1:

One kWh of electricity emits 0.85 pounds of CO2. Assuming that the interim SCC correctly reflects the total social cost of one metric ton of CO2, what is the marginal externality cost per kWh of electricity?

```{r}
# interim SCC is $51 per metric ton CO2
scc_int <- 51

# convert from tons to lbs 
scc_int_lbs <- scc_int / 2204.62

# we are told that 1kWh produces 0.85 lbs CO2. Use this to calculate SCC per kWh
mec <- scc_int_lbs * 0.85

```

Answer: The marginal externality cost is $`r round(mec, 4)` per kWh of electricity.


### Q2: 
What is the aggregate monthly demand curve for electricity? What is the supply curve for electricity? What is the “benefit” to consumers under the status quo? What is the “benefit” to producers under the status quo? What is the environmental cost under the status quo?


```{r}
# Linear regression models

model_demand_l <- lm(price_cents ~ q_low_kwh, data = e_data)
model_demand_h <- lm(price_cents ~ q_high_kwh, data = e_data)

low_tidy <- tidy(model_demand_l)
low_tidy

high_tidy <- tidy(model_demand_h)
high_tidy
```

need to rearrange the parameter to get Q(P)! 

Qgg = Qlow(P) + Qlow(h) 

Importantly, since they-intercepts are different, we know that Qagg(P) will have a kink. I include an ifelse() statement to take care of the kink.

define a function to get demand

```{r}
# need to rearrange the parameter to get Q(P)! 

# Qgg = Qlow(P) + Qlow(h) 

# Importantly, since they-intercepts are different, we know that Qagg(P) will have a kink. I include an ifelse() statement to take care of the kink.

# define a function to get demand

demand <- function(p, model){
  q <- (p - model$coefficients[[1]])/model$coefficients[[2]]
  q <- ifelse(q<0,0,q)
  return(q)
}

# for each p level, return aggregate demand
demand_agg <- function(p){
  q <- demand(p, model_demand_l) + demand(p, model_demand_h)
  return(q)
}

# trying to replicate for low-income demand
demand_low <- function(p){
  q <- demand(p, model_demand_l)
  return(q)
}

# trying to replicate for high-income demand
demand_high <- function(p){
  q <- demand(p, model_demand_h)
  return(q)
}

# current market price is $0.10 per kWh
market_q <- demand_agg(10)
```

```{r}
# Sandy's code for Qagg
price = seq(0, 31.6, length.out = 100)
Qagg <- map(price, demand_agg) %>% unlist()

df<- tibble(Qagg = Qagg, price = price)

# trying to do the same for Qlow
Qlow <- map(price, demand_low) %>% unlist()
Qlow_df <- tibble(Qlow = Qlow, price = price)

# trying to do same for Qhigh
Qhigh <- map(price, demand_high) %>% unlist()
Qhigh_df <- tibble(Qhigh = Qhigh, price = price)

```


```{r}
# plot the whole thing!
ggplot() +
  # graph the agg demand
  geom_line(data = df, aes(Qagg, price), color = "black", size = 1.1) +
  # graph low demand
  geom_line(data = Qlow_df, aes(Qlow, price), color = "slateblue", size = 1, linetype = "twodash") +
  # graph high demand
  geom_line(data = Qhigh_df, aes(Qhigh, price), color = "sienna", size = 1, linetype = "dashed") +
  # graph current MPC
  geom_abline(slope = (10/market_q), intercept = 0, color = "purple", size = 1)+
  
  # graph the low income points
  geom_point(data = e_data, aes(x = q_low_kwh, y = price_cents), color = "slateblue", alpha = 0.5, size = 2) +
  # graph the high income points
  geom_point(data = e_data, aes(x = q_high_kwh, y = price_cents), color = "sienna", alpha = 0.5, size = 2) +

  labs(x = "\nQuantity (kWh)",
       y = "Price (cents)\n") +
  
  theme_classic() +
  
  theme(
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold")) +
  # force origin at 0,0
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0))

```

The marginal private cost (supply) curve for electricity is: 
$MPC= 0.1Q $

```{r}
# I also define functions for calculating the consumer surplus:

CS <- function(p, model){
  q <- demand(p, model)
  cs <- 0.5*(model$coefficients[[1]] - p)*q
  return(cs)
}

CS_agg <- function(p){
  cs <- CS(p,model_demand_l) + CS(p,model_demand_h)
  return(cs)
}
```


```{r}
# calculating surpluses at market equilibrium
# CS calculation needs Y intercept of high demand/demand agg curve --> at 0 kWh, price would be $0.316
high_sum <-summary.lm(model_demand_h)
high_y_int <- high_sum$coefficients[1, 1] 
low_sum <- summary.lm(model_demand_l)
low_y_int <- low_sum$coefficients[1, 1]
# Converted back to dollars
cs_sq_high <- 0.5*market_q*((high_y_int/100) - 0.10)
# High consumer benefit $57,984.10
cs_sq_low <- 0.5*market_q*((low_y_int/100) - 0.10)
# Low consumer benefit $35,882.29
cs_sq_total <- cs_sq_high + cs_sq_low
# Consumer benefit $93,866.39
ps_sq <- 0.5*market_q*0.10
# Producer benefit $26,835.97
env_cost_sq <- (MEC*market_q)/100
# Environmental cost $10,553.65
```

The benefit, or surplus, to consumers 


































```{r}
ggplot(data = e_data) +
  # graph the low income demand
  geom_point(aes(x = q_low_kwh, y = price_cents), color = "slateblue", alpha = 0.5, size = 2) +
  
  # graph the high income demand
  geom_point(aes(x = q_high_kwh, y = price_cents), color = "sienna", alpha = 0.5, size = 2) +

    labs(x = "\nQuantity (kWh)",
       y = "Price ($/kWh)\n") +
  
  theme_classic() +
  
  theme(
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold")) 
```
